Interfacing logic programming and Coq

### 2020-10-26 papers/types2020/cultures.tex
Changing machine

Proof search has proved to have some richness that is hard to capture
in the proof normalization setting.  In particular, Gentzen's
invention of \emph{eigenvariables} are a kind of a proof-level
binder.  In the proof normalization setting, such eigenvariables
are instantiated during proof normalization.  In contrast, during the
search for cut-free proofs, eigenvariables are part of the syntax of
terms and formulas.  As a result, they can be used in the treatment of
bindings in data structures more generally.  Hence, the logic
programming paradigm provides a completely natural, elegant, and
powerful method, called $\lambda$-tree syntax, for treating bindings
within syntax.

@@

Given that \lP and Coq are both examples of mixing the
$\lambda$-calculus with logic, it is important to understand their
differences.  The confusion around the term \emph{higher-order
abstract syntax} (HOAS), for example, helps to illustrate their
differences.  In the functional programming setting, including the Coq
system, the HOAS approach leads to encoding binding structures within
terms using functions.  Such encodings have results in surprisingly
complicated encodings, allowing, for example, for \emph{exotic terms}
\cite{despeyroux95tlca} and to structures on which induction is not
possible \cite{roeckl01fossacs}.  In the logic programming setting,
particularly \lP, HOAS is well supported in both the availability of
binders allowed with terms ($\lambda$-bindings), in formulas
(quantifiers), and in proofs (eigenvariables).  For this reason, the
term $\lambda$-tree syntax was introduced to name this particular take
on HOAS \cite{miller19jar}.  The Abella proof assistant
\cite{baelde14jfr} was designed, in part, to allow for inductive and
co-inductive proofs for specifications using the $\lambda$-tree syntax
approach.

>>>>

Proof search has proved to have some richness that is hard to capture
in the proof normalization setting.  In particular, Gentzen's
invention of \emph{eigenvariables} are a kind of a proof-level
binder.  In the proof normalization setting, such eigenvariables
are instantiated during proof normalization.  In contrast, during the
search for cut-free proofs, eigenvariables are part of the syntax of
terms and formulas.  As a result, they can be used in the treatment of
bindings in data structures more generally.  Hence, the logic
programming paradigm provides a completely natural, elegant, and
powerful method for treating bindings
within syntax.

@@

Given that \lP and Coq are both examples of mixing the
$\lambda$-calculus with logic, it is important to understand their
differences.  The confusion around the term \emph{higher-order
abstract syntax} (HOAS), for example, helps to illustrate their
differences.  In the functional programming setting, including the Coq
system, the HOAS approach leads to encoding binding structures within
terms using functions.  Such encodings have results in surprisingly
complicated encodings, allowing, for example, for \emph{exotic terms}
\cite{despeyroux95tlca} and to structures on which induction is not
possible \cite{roeckl01fossacs}.  In the logic programming setting,
particularly \lP, HOAS is well supported in both the availability of
binders allowed with terms ($\lambda$-bindings), in formulas
(quantifiers), and in proofs (eigenvariables).  For this reason, the
term \emph{$\lambda$-tree syntax} was introduced to name this
particular take on HOAS \cite{miller19jar}.  The Abella proof
assistant \cite{baelde14jfr} was designed, in part, to allow for
inductive and co-inductive proofs for specifications using the
$\lambda$-tree syntax approach.
